Convergence rate of MAP estimates \\ for the exponential family

### 2021-10-05 main.tex
Add new drawings to manuscript

\paragraph{Duality}
The logpartition function $\logpart$ verifies the two following identities
\begin{align}
    \nabla\logpart(\natp) &=  \expect[p(X|\natp)]{T(X)} =: \meanp \\
    \nabla^2 \logpart(\natp) &= \Cov_\natp[T(X)] > 0
\end{align}
where $\meanp$ is called the mean parameter.
If the sufficient statistic $T$ is minimal, then the log-partition function $\logpart$ is strictly convex and its gradient $\nabla \logpart$ is a bijection between natural parameters $\natp$ and mean parameters $\mu$.
The second identity entails that $\logpart$ is strictly-convex. 
At this point it is useful to introduce the \href{https://en.wikipedia.org/wiki/Convex_conjugate}{convex conjugate} (aka Fenchel-Legendre transform) of the logpartition function
\begin{align}
	\conj(\mu) = \langle \mu, \natp \rangle - \logpart(\natp) \; .
\end{align}
It turns out that $\conj$ matches the common notion of \textit{entropy} in information theory, so we will call it entropy.
If $\logpart$ is strictly convex, then its gradient is strictly monotone, so it is a bijection, and its inverse is the gradient of its dual $\nabla\conj \circ \nabla\logpart(\natp) = \natp$ (cf Fig.~\ref{fig:duality}).
For a full review of exponential families and their duality, see \citet[Chapter 3]{wainwright2008graphical}.
\begin{figure}[ht]
	\centering
	\includegraphics{figs/duality}
	\caption{The gradient of the log-partition function and its dual, $(\nabla \logpart, \nabla \conj)$, form a bijection between the natural and mean parameters $\natp, \meanp$. Figure reproduced from \citet{kunstner2020homeomorphic}.}
	\label{fig:duality}
\end{figure}
\begin{example}
For $\cN(0,\sigma^2)$,we can use the formula $\conj(\mu) = \mu \natp - \logpart(\natp)$ to get the entropy
\begin{align}
	\conj(\mu) = \half\left( -\log(\mu) + \log\frac{\pi}{2} - 1 \right) \; .
\end{align}
We can also take gradient and derivative of $\logpart(\natp)$ to retrieve the mean and covariance of the sufficient statistic $X^2$
\begin{align}
	\nabla\logpart(\natp) &= \frac{-1}{2 \natp} = \sigma^2 = \mu = \E[X^2] \\
	\nabla^2\logpart(\natp) &= \frac{1}{2 \natp^2} = 2 \sigma^4 = 2 \mu^2 = \Var(X^2) 
\end{align}
which we confirm thank to wikipedia since $\E[X^4] = 3 \sigma^4$ and thus $\Var(X^2) = \E[X^4] - \E[X^2]^2 = 3 \sigma^4 - \sigma^4 = 2 \sigma^4$.
\end{example}

@@

\begin{figure}[ht]
	\centering
	\includegraphics[width=\textwidth]{figs/thales/numerical_schema_n=3.pdf}
	\caption{A numerical illustration of the different characters featured in the bias-variance decomposition, for a 1D Gaussian $\cN(\mu, \sigma^2)$.}
	\label{fig:bias-variance-numerical}
\end{figure}

>>>>

\paragraph{Duality}
The logpartition function $\logpart$ verifies the two following identities
\begin{align}
    \nabla\logpart(\natp) &=  \expect[p(X|\natp)]{T(X)} =: \meanp \\
    \nabla^2 \logpart(\natp) &= \Cov_\natp[T(X)] > 0
\end{align}
where $\meanp$ is called the mean parameter.
If the sufficient statistic $T$ is minimal, then the log-partition function $\logpart$ is strictly convex and its gradient $\nabla \logpart$ is a bijection between natural parameters $\natp$ and mean parameters $\mu$.
The second identity entails that $\logpart$ is strictly-convex. 
At this point it is useful to introduce the \href{https://en.wikipedia.org/wiki/Convex_conjugate}{convex conjugate} (aka Fenchel-Legendre transform) of the logpartition function
\begin{align}
	\conj(\mu) = \langle \mu, \natp \rangle - \logpart(\natp) \; .
\end{align}
It turns out that $\conj$ matches the common notion of \textit{entropy} in information theory, so we will call it entropy.
If $\logpart$ is strictly convex, then its gradient is strictly monotone, so it is a bijection, and its inverse is the gradient of its dual $\nabla\conj \circ \nabla\logpart(\natp) = \natp$ (cf Fig.~\ref{fig:duality}).
For a full review of exponential families and their duality, see \citet[Chapter 3]{wainwright2008graphical}.
\begin{figure}[ht]
	\centering
	\includegraphics[width= .6\textwidth]{figs/duality} \\
	\includegraphics[width=.7\textwidth]{figs/2d/mirrormap.pdf}
	\caption{The gradient of the log-partition function and its dual, $(\nabla \logpart, \nabla \conj)$, form a bijection between the natural and mean parameters $\natp, \meanp$.Top figure reproduced from \citet{kunstner2020homeomorphic}. Bottom figure represents $\cN(\mu, \sigma^2)$.}
	\label{fig:duality}
\end{figure}
\begin{example}
For $\cN(0,\sigma^2)$,we can use the formula $\conj(\mu) = \mu \natp - \logpart(\natp)$ to get the entropy
\begin{align}
	\conj(\mu) = \half\left( -\log(\mu) + \log\frac{\pi}{2} - 1 \right) \; .
\end{align}
We can also take gradient and derivative of $\logpart(\natp)$ to retrieve the mean and covariance of the sufficient statistic $X^2$
\begin{align}
	\nabla\logpart(\natp) &= \frac{-1}{2 \natp} = \sigma^2 = \mu = \E[X^2] \\
	\nabla^2\logpart(\natp) &= \frac{1}{2 \natp^2} = 2 \sigma^4 = 2 \mu^2 = \Var(X^2) 
\end{align}
which we confirm thank to wikipedia since $\E[X^4] = 3 \sigma^4$ and thus $\Var(X^2) = \E[X^4] - \E[X^2]^2 = 3 \sigma^4 - \sigma^4 = 2 \sigma^4$.
\end{example}

@@

\begin{figure}[ht]
	\centering
	\includegraphics[width=\textwidth]{figs/thales/numerical_schema_n=3.pdf}
	\includegraphics[width=\textwidth]{figs/2d/gaussian_convergence.pdf}
	\caption{A numerical illustration of the different characters featured in the bias-variance decomposition, for a 1D Gaussian $\cN(\mu, \sigma^2)$, for 1 time step  and many repetitions with $n=3$ (top) illustrating the effect of $n_0$ with the Thales triangle. Or $50$ time-steps at once and 1 repetition (bottom), illustrating the convergence behavior.}
	\label{fig:bias-variance-numerical}
\end{figure}
